\documentclass[12pt]{article}

\usepackage{tikz} % картинки в tikz
\usepackage{microtype} % свешивание пунктуации

\usepackage{array} % для столбцов фиксированной ширины

\usepackage{indentfirst} % отступ в первом параграфе

\usepackage{sectsty} % для центрирования названий частей
\allsectionsfont{\centering}

\usepackage{amsmath, amsfonts, amssymb} % куча стандартных математических плюшек

\usepackage{comment}

\usepackage[top=2cm, left=1.2cm, right=1.2cm, bottom=2cm]{geometry} % размер текста на странице

\usepackage{lastpage} % чтобы узнать номер последней страницы

\usepackage{enumitem} % дополнительные плюшки для списков
%  например \begin{enumerate}[resume] позволяет продолжить нумерацию в новом списке
\usepackage{caption}


\usepackage{fancyhdr} % весёлые колонтитулы
\pagestyle{fancy}
\lhead{Time Series and Stochastic Processes}
\chead{}
\rhead{HA}
\lfoot{2020-2021}
\cfoot{}
\rfoot{\thepage/\pageref{LastPage}}
\renewcommand{\headrulewidth}{0.4pt}
\renewcommand{\footrulewidth}{0.4pt}



\usepackage{todonotes} % для вставки в документ заметок о том, что осталось сделать
% \todo{Здесь надо коэффициенты исправить}
% \missingfigure{Здесь будет Последний день Помпеи}
% \listoftodos - печатает все поставленные \todo'шки


% более красивые таблицы
\usepackage{booktabs}
% заповеди из докупентации:
% 1. Не используйте вертикальные линни
% 2. Не используйте двойные линии
% 3. Единицы измерения - в шапку таблицы
% 4. Не сокращайте .1 вместо 0.1
% 5. Повторяющееся значение повторяйте, а не говорите "то же"



\usepackage{fontspec}
\usepackage{polyglossia}

\setmainlanguage{english}
\setotherlanguages{english}

% download "Linux Libertine" fonts:
% http://www.linuxlibertine.org/index.php?id=91&L=1
\setmainfont{Linux Libertine O} % or Helvetica, Arial, Cambria
% why do we need \newfontfamily:
% http://tex.stackexchange.com/questions/91507/
\newfontfamily{\cyrillicfonttt}{Linux Libertine O}

%\AddEnumerateCounter{\asbuk}{\russian@alph}{щ} % для списков с русскими буквами
%\setlist[enumerate, 2]{label=\asbuk*),ref=\asbuk*}

%% эконометрические сокращения
\DeclareMathOperator{\Cov}{\mathbb{C}ov}
\DeclareMathOperator{\Corr}{\mathbb{C}orr}
\DeclareMathOperator{\Var}{\mathbb{V}ar}
\DeclareMathOperator{\E}{\mathbb{E}}
\DeclareMathOperator{\tr}{trace}
\DeclareMathOperator{\card}{card}

\newcommand \hb{\hat{\beta}}
\newcommand \hs{\hat{\sigma}}
\newcommand \htheta{\hat{\theta}}
\newcommand \s{\sigma}
\newcommand \hy{\hat{y}}
\newcommand \hY{\hat{Y}}
\newcommand \e{\varepsilon}
\newcommand \he{\hat{\e}}
\newcommand \z{z}
\newcommand \hVar{\widehat{\Var}}
\newcommand \hCorr{\widehat{\Corr}}
\newcommand \hCov{\widehat{\Cov}}
\newcommand \cN{\mathcal{N}}
\newcommand \RR{\mathbb{R}}
\newcommand \NN{\mathbb{N}}
\newcommand{\cF}{\mathcal{F}}
\newcommand{\cH}{\mathcal{H}}


\begin{document}

\section*{Home Assignment 1}

\begin{enumerate}
  \item Let $\Omega = \RR$. Explicitely find the sigma-algebras $\cF_1 = \sigma(A)$, $\cF_2 = \sigma(B)$, $\cF_3 = \sigma(A, B)$ where $A=[-10;5]$ and $B=(0;10)$.
  \item I throw a die once. Let $X$ be the result of the toss. 
  Count the number of events in sigma-algebras $\cF_1 =\sigma (X)$, $\cF_2 = \sigma (\{X>3\})$, $\cF_3 = \sigma (\{X > 3\}, \{X<5\})$.

  \item Let $\Omega = \RR$. The sigma-algebra $\cF$ is generated by all the sets of the form $(-\infty, t]$,
  \[
  \cF = \sigma \left( \{ (-\infty; t] \mid t \in \RR\} \right)  
  \]
  Check whether $A_1 = (0; 10) \in \cF$, $A_2 = \{5\} \in \cF$, $A_3 = \NN \in \cF$. 

  \item Prove the following statements or provide a counter-example:
  \begin{enumerate}
    \item If $\cF_1$ and $\cF_2$ are sigma-algebras then $\cF = \cF_1 \cup \cF_2$ is sigma-algebra.
    \item If $X$ and $Y$ are independent random variables then $\card \sigma(X, Y) = \card \sigma(X) + \card \sigma(Y)$.
  \end{enumerate}

  For finite sets $\card$ denotes just the number of elements.

  \item I throw a die infinite number of times. 
  Let the random variable $X_n$ be equal to $1$ if the $n$-th toss is head and $0$ otherwise. 
  Consider a pack of sigma-algebras: $\cF_n = \sigma (X_1, \ldots, X_n)$ and $\cH_n = \sigma (X_n, X_{n+1}, X_{n+2}\, \ldots)$.

  Where possible provide and example of a non-trivial event (neither $\Omega$ nor $\emptyset$) such that 
  \begin{enumerate}
    \item $A_1 \in \cF_{2020}$;
    \item $A_2 \in \cH_{2020}$;
    \item $A_3 \in \cF_{2020}$ and $A_3 \in \cH_{2020}$;
    \item $A_4 \in \cF_n$ for all $n$; 
    \item $A_5 \in \cH_n$ for all $n$.
  \end{enumerate}

\end{enumerate}


Deadline: 25 September 2020, 21:00 MSK.


\newpage
\section*{Home Assignment 2}
\begin{enumerate}

  \item Consider the Markov chain with the transition matrix:
  
  \[
    \begin{pmatrix}
      0   & 0.1 & 0.9 &   0 &   0 &   0 &   0  &  0 \\
      0.5 & 0.5 & 0   &   0 &   0 &   0 &   0  &  0 \\
      0   & 0   & 0   &   0.2 &   0 &   0 &   0  &  0.8 \\
      0   & 0 & 0.3 &   0 &   0.7 &   0 &   0  &  0 \\
      0   & 0 & 0 &   0 &   0 &   1 &   0  &  0 \\
      0   & 0 & 0 &   0 &   0 &   0 &   1  &  0 \\
      0   & 0 & 0 &   0 &   0 &   0.4 &   0  &  0.6 \\
      0   & 0 & 0 &   0 &   0 &   1 &   0  &  0 \\
    \end{pmatrix}  
  \]
  
  
  \begin{enumerate}
    \item Draw the most beautiful graph for this chain. A fox or a cat is ok :)
    \item Classify the states of the Markov chain.
  \end{enumerate}
  
  


  \item The Lonely Knight is standing on the A1 field of the chessboard. 
  She starts moving randomly according to chess rules. 

  \begin{enumerate}
    \item How many moves on average will it take to go back to A1?
    \item What proportion of her eternal life will she spend on every field?
  \end{enumerate}

  \item Donald Trump throws a die until one appears or until he says «Stop».
  The payoff is equal to the last thrown number. Donald maximizes the expected payoff. 

  
  \begin{enumerate}
    \item What is the best strategy and the corresponding expected payoff?
  \end{enumerate}

  How do the answers change in the following modifications of the orignal game?

  \begin{enumerate}[resume]
    \item Donald is also required to stop at 3 and to continue on 4.
    \item Donald should pay 0.3 for every throw.
  \end{enumerate}

\item Ilya Muromets stands before the first stone. There are three roads behind the stone. 
And every road ends with a new stone. And there are three new roads behind every new stone. And so on. 
Every road is guarded with one-third probability by a three headed Zmei Gorynich.
Yes, there are infinitely many Zmei Gorynich.

What is the probability that there is an Eternal Peaceful Journey?

\item Ilya and Zmei finally met and play with a coin. 
They throw a coin until the sequence HTT or TTH apears. 
Ilya wins if HTT appears and Zmei wins if TTH appears.

\begin{enumerate}
  \item What is the probability that Ilya wins?
  \item What is the expected number of throws?
  \item What is the expected number of throws given that Ilya won?
\end{enumerate}

\end{enumerate}




Deadline: 9 October 2020, 21:00 MSK.


\newpage
\section*{Home Assignment 2}
\begin{enumerate}
  \item 
\end{enumerate}


Deadline: 25 September 2020, 21:00 MSK.


\end{document}
