\documentclass[12pt]{article}

\usepackage{tikz} % картинки в tikz
\usepackage{microtype} % свешивание пунктуации

\usepackage{array} % для столбцов фиксированной ширины

\usepackage{indentfirst} % отступ в первом параграфе

\usepackage{sectsty} % для центрирования названий частей
\allsectionsfont{\centering}

\usepackage{amsmath, amssymb, amsthm} % куча стандартных математических плюшек

\usepackage{amsfonts}

\usepackage{comment}

\usepackage[top=2cm, left=1.2cm, right=1.2cm, bottom=2cm]{geometry} % размер текста на странице

\usepackage{lastpage} % чтобы узнать номер последней страницы

\usepackage{enumitem} % дополнительные плюшки для списков
%  например \begin{enumerate}[resume] позволяет продолжить нумерацию в новом списке
\usepackage{caption}


\usepackage{hyperref} % гиперссылки

\usepackage{multicol} % текст в несколько столбцов


\usepackage{fancyhdr} % весёлые колонтитулы
\pagestyle{fancy}
\lhead{Time Series and Stochastic Processes}
\chead{Midterm exam 2020}
\rhead{2020-10-30}
\lfoot{}
\cfoot{}
\rfoot{}
\renewcommand{\headrulewidth}{0.4pt}
\renewcommand{\footrulewidth}{0.4pt}

\let\P\relax
\DeclareMathOperator{\P}{\mathbb{P}}
\DeclareMathOperator{\E}{\mathbb{E}}
\DeclareMathOperator{\Var}{\mathbb{V}ar}

\usepackage{todonotes} % для вставки в документ заметок о том, что осталось сделать
% \todo{Здесь надо коэффициенты исправить}
% \missingfigure{Здесь будет Последний день Помпеи}
% \listoftodos - печатает все поставленные \todo'шки


% более красивые таблицы
\usepackage{booktabs}
% заповеди из докупентации:
% 1. Не используйте вертикальные линни
% 2. Не используйте двойные линии
% 3. Единицы измерения - в шапку таблицы
% 4. Не сокращайте .1 вместо 0.1
% 5. Повторяющееся значение повторяйте, а не говорите "то же"



\usepackage{fontspec}
\usepackage{polyglossia}

\setmainlanguage{english}
\setotherlanguages{russian}

% download "Linux Libertine" fonts:
% http://www.linuxlibertine.org/index.php?id=91&L=1
\setmainfont{Linux Libertine O} % or Helvetica, Arial, Cambria
% why do we need \newfontfamily:
% http://tex.stackexchange.com/questions/91507/
\newfontfamily{\cyrillicfonttt}{Linux Libertine O}

\AddEnumerateCounter{\asbuk}{\russian@alph}{щ} % для списков с русскими буквами
% \setlist[enumerate, 2]{label=\asbuk*),ref=\asbuk*}




\begin{document}

Here $(W_t)$ denotes the standard Wiener process.

\begin{enumerate}
    
    
    
    \item For $r<s<t<u$ find the following expected values 
    \begin{enumerate}
    \item $\E((W_u - W_t)^2(W_s - W_r)^2)$;
    \item $\E((W_u - W_s)(W_t - W_r))$;
    \item $\E((W_t - W_r)(W_s - W_r)^2)$;
    \item $\E(W_r W_s W_t)$;
    \item $\E(W_r W_s W_t \mid W_s)$;
    \end{enumerate}

\item Consider Ito process $X_t$

\[
dX_t = \exp(t) W_t\, dt + \exp(2W_t) \, dW_t, \quad X_0 = 1.
\]

Consider two processes, $A_t = 1 + t^2 + X_t^3$ and $B_t = 1 + t^2 + X_t^3 W_t^4$.

\begin{enumerate}
    \item Find $dA_t$ and $dB_t$.
    \item Write the corresponding explicit expressions for $A_t$ and $B_t$:
    \textcolor{red}{
    \[
    const + \int_0^t \ldots dW_u + \int_0^t \ldots du
    \]
    }
    \item Check whether $X_t$ is a martingale.
\end{enumerate}

\item Let $S_0 = 0$, $S_t = X_1 + X_2 + \ldots + X_t$. The increments $X_t$ are independent and identically distributed: 

\begin{tabular}{cccc}
\toprule
$x$ & $-1$ & $0$ & $1$ \\
$\P(X_t = x)$ & $0.2$ & $0.2$ & $0.6$ \\
\bottomrule
\end{tabular}

\begin{enumerate}
    \item If possible find all constants $a$ such that $M_t = S_t + at$ is a martingale.
  \item If possible find all constants $b$ such that $R_t = b^{S_t}$ is a martingale.
\end{enumerate}

\item Consider the process $X_t$

\[
X_t= tW_t + \int_0^t uW_u^2\, dW_u.
\]

\begin{enumerate}
    \item Find $\E(X_t)$, $\Var(X_t)$.
    \item Find $dX_t$.
    \item Check whether $X_t$ is a martingale.
\end{enumerate}

\item A Hedgehog in the fog starts in $(0, 0)$ at $t=0$ and moves randomly with equal probabilities in four directions (north, south, east, west) by one unit every minute. 

Let $X_t$ and $Y_t$ be his coordinates after $t$ minutes \textcolor{red}{and $S_t = X_t + Y_t$}.

\begin{enumerate}
    \item Find $\E(X_2 \mid S_2)$;
    \item Find $\Var(X_2 \mid S_2)$.
\end{enumerate}

Hint: $\Var(Y \mid X) = \E(Y^2 \mid X) - (\E(Y \mid X))^2$.

\newpage
    \item Vampire Petr and Markov Chains. 
    
    Vampire Petr drinks blood of a new victim every day. 
    Unfortunately 20\% of the population are vaccinated against vampires. 
    If more than one victim of the last three victims are vaccinated then Petr will be instantaneously cured and will return to the normal life. 
 
    For simplicity let's assume that the last three victims were not vaccinated. 
    
    \begin{enumerate}
        \item What is the probability that vampire Petr will be cured in the next three days?
        \item How many victims will be bitten by vampire Petr on average?
    \end{enumerate}
 
    \item Vampire Boris and Martingales.
    
    To survive vampire Boris needs to bite 70 talented students. 
    
    These 70 talented students have formed a secret group. They have written their emails on small pieces of paper and have randomly distributed these pieces among them. Each student has exactly one piece of paper with an email\footnote{The group is so secret that it is possible that a student has his own email on his piece of paper}. 
    
    Initially vampire Boris knows contacts of just two persons from the group. Today he will contact them, drink their blood and get the emails they have. Then vampire Boris will contact new victims and so on.
    
    \begin{enumerate}
        \item \textcolor{red}{For $t\geq 1$} consider the process $M_t$, the proportion of non bitten students after the day $t$. 
        
        Is this process a martingale?
        
        \item Using martingale stopping theorem or otherwise find the probability that vampire Boris will bite all 70 students. 
    \end{enumerate}
 
 
\end{enumerate}


\end{document}
